---
title: "Intro to Categorical Variables"
author: "P.Lombardo"
output:
  html_document:
    df_print: paged
---

Before beginning, let's get used to loading the `R` packages we might need for analysis later.
```{r}
library(tidyverse)
```
The `tidyverse` is actually a collection of packages that all work well together, and it includes functions for data wrangling, plotting, and much more!  We will learn to use the tidyverse well this semester.

# Exploring and Summarizing Categorical Variables

Let's work with our titanic data set again.  Can you remember how to load this CSV file as a data frame in R?
```{r}
# Place your code here, or use the import data set button.
```

The `table()` command is a quick way to have `R` create tables of counts for categorical variables.  For example, the `Survived` variable in our titantic data set is a categorical variable (a "factor") with two "levels":

  * Yes (indicating that person survived); and
  * No (indicating that person died).

Run the following command to count how many people survived.
```{r}
table(titanic$Survived)
```

How many survived?
<Place your answer here>

#### Exercise 1:
How many females were on board the titanic according to this data set?
```{r}
#place your code here.
```


#### Exercise 2:
What *percentage* of the passengers on board were male?
```{r}
#place your code here.

```


#### Exercise 3:
Let's look into using `prop.table()` and the "piping operator", `%>%`.
```{r}
# place your code here.

```


****

## Contingency tables for *two* categorical variables

Can we find statistical evidence for the phrase?

>``Women (and children) First."

This involves exploring...

* What would "evidence" involve? What variables do we need to look at? How should we summarize these variables?
* If the evidence is in favor of the claim, what would it look like? What specifically would see in our summaries?


We can create "contingency tables" (or cross tabs, there are many names) by passing two variables into the `table()` command.  Run the following code:
```{r}
table(titanic$Gender, titanic$Survived)

# or the "piped" approach
titanic %>%
    select(Gender, Survived) %>%
    table()
```

*Note*: whichever variable is "selected" first will appear as the rows, the second variable appears as the columns.


#### Exercise 3:
Explain what each number in the table counts; be specific and describe how the row and column labels tell you this information.

<place your answer here>

****

#### Exercise 4:
Using `R` as a calculator, use the numbers in the table above to compute the survival rate for men and the survival rate for women. 

*Note*: the survival rate for men would be the percentage *of the males* that survived. 
```{r}
#place your code here
```

Which `Sex` had the higher survival rate?
<place your answer here>

****

We can compute row or column percentages quickly by passing a table into the `prop.table()` command in choosing a `margin`.

Try this code:
```{r}
titanic %>% 
    select(Gender, Survived) %>% 
    table() %>%
    prop.table(margin =1)
```

#### Exercise 5:
Answer the following questions:

* What happens if you set `margin = 2`?
    * What happens if you do not specify a "margin" at all?

* Using a similar command as above, find the percentage of those that survived who were female.

```{r}
# place your code here.
```


