---
title: "Visualizing Categorical Data"
author: "P.Lombardo"
output:
  html_document:
    df_print: paged
---

Before beginning, let's get used to loading the `R` packages we might need for analysis later.
```{r}
library(tidyverse)
titanic <-read.csv("data/titanic.csv",
                   header = T,
                   stringsAsFactors = T)
```

## Visualizing frequencies and percentages
When working with categorical data, there are many visualizations. However, the most basic visual is a bar plot (and some of its varieties), which you have read about already.  I also happen to believe that the most basic plot is also the best! We have:

* Simple bar plot
* Side-by-side bar plots
* Stacked bar plots
* 100\%-stacked bar plots

Now let's learn how to make bar plots using the visualization package `ggplot`.

## Pretty visuals with `ggplot2`
The plotting package `ggplot2` (part of the `tidyverse`) uses a layering approach to building a plot. Let's build a standard bar plot together, and we'll talk about each layer.

```{r}
# place code here

```

### Visuals with two variables
When visualizing two variables, we have three options: a side-by-side barplot, a stacked barplot, and a 100%-stacked bar plot.

* For each of these plots, there is a "main grouping variable" and its levels appear as the labels for the $x$-axis.
* The second variable affects how we *fill in* the bars we just created with color. In ggplot, we use the "fill" aesthetic to control this.

#### Side-by-side barplot
The main idea is to set the main grouping variable as the `x` aesthetic, and the secondary variable as the `fill` aesthetic.  To get the side-by-side-ness, we pass `position = "dodge"` inside the `geom_bar()` layer.

Since we are interested in how `Gender` associates with `Survived`, we want male and female to be considered separately on our horizontal axis. Hence, `Gender` is our main grouping variable, and `Survived` is the secondary variable.

**Exercise.** Fill in the dots below with the correct variable names and run the code.

```{r}
ggplot(data = titanic,
       aes(x = ..., fill = ...))+
    geom_bar(position = "dodge")
```

#### Stacked barplot
This is identical to the plot above, but drop the `position="dodge"` command from `geom_bar()`.  

**Exercise.** Copy and paste your code above, and remove the appropriate argument from `geom_bar()`. Run your code and see if it works!
```{r}
# Place your copy here and modify it.
```


#### 100%-stacked barplot
Again, this is a small modification on the previous plots.  Place the argument `position="fill"` into the `geom_bar()` layer.

**Exercise.** Copy and paste your code above, and make the necessary modification.

```{r}
# Place your copy here and modify it.
```

### Important reflection!
Each of the plots we created above highlight a different component of the association between `Gender` and `Survived` for the titanic disaster.  With a partner, write a few sentences on the strengths and weaknesses of each kind of barplot (side-by-side, stacked, and 100%-stacked)

<Place your paragraph here.>

***


## (Optional) Modifying ggplots
### Can we change the way a ggplot looks?
Yes! You can do pretty much anything.  Below is a bit of sample code that makes some common tweaks to the plot. I've left comments to help you loosely understand it.

*Note*: You do not have to understand all the code below. I provide this more to let you know some of the things ggplot can adjust.
```{r}
ggplot(data = titanic,
       aes(x = Pclass,
           fill = Survived))+
    geom_bar(position = "fill")+
    
    # pick a new color scheme by "scaling" your "fill" aesthetic
    scale_fill_brewer("Survived (Yes/No)", # changes the legend title
                      palette="Set1" # brewer palettes are premade
                      ) +
    
    # Adjust labels and create titles, subtitles, captions
    labs(x = "Passenger Class", #changes x-axis label
         y = "Percentages", #changes y-axis label
         title = "Comparing Survival Rates by Passenger Class",
         subtitle = "Titanic Disaster",
         caption = "(And this is why I won't go on a cruise... I can't afford first class.)"
         ) + 
    
    # make the y axis look "nicer"
    scale_y_continuous(breaks = seq(0,1,by=.1), # show every 10%
                       labels = scales::percent # format percents
                       ) +
    
    # remove gray background
    theme_bw()
```

### Mosaic plots
Another common plot is a mosaic plot.
```{r}
titanic %>% with(
    table(Gender, Survived)
) %>% mosaicplot(color=T)
```

We can generate these with ggplot if we add a new package: `ggmosaic`. I will skip the details, but you can read about them [here](https://cran.r-project.org/web/packages/ggmosaic/vignettes/ggmosaic.html)