---
title: "Exploring Berkeley Admissions Data"
subtitle: "Practice with Categorical Variables"
author: "P. Lombardo"
output:
  html_document:
    df_print: paged
---

## Loading packages and data
```{r}
library(tidyverse)
berkeley<-read.csv('data/berkeley.csv', header = T,
              stringsAsFactors = T)
```

## Getting started with statistical thinking
Above we loaded the Berkeley data set, which looks at acceptance rates to programs in 1973. 

How many variables are there in this data set? What are they?
```{r}
#place code here
```

The original intent of this data was to explore whether there was a gender bias in Berkeley college admissions at that time.  What initial summary helps us answer this question?

Compute this summary below and create a visual for it:
```{r}
#place code here
```

Often in *observational studies*, where there is no experimental assignment of a treatment, a "confounding variable" can have an influencing effect on the research question, making a straightforward analysis like the one we did above not really capture the nuance of the situation.

With this in mind, create a summary table that compares acceptance rates for males and females, ***but broken up by `Major`***.  As usual, provide an appropriate visual as well.
```{r}
#place code here
```

What do you notice?  Something surprising?
```{r}
# place code here
```

This is an example of something called [Simpson's Paradox](https://en.wikipedia.org/wiki/Simpson%27s_paradox), where there is a clear trend when data is considered by groups separately, but that trend *reverses or disappears* when they groups are aggregated and treated as one.

Tricky, tricky, tricky....

Let's use the remaining time to get started on our lab for the weekend!
