---
title: "Percentiles and Boxplots"
author: "P. Lombardo"
output:
  html_document: default
---

## Loading packages and data
```{r}
library(tidyverse)
small_diamonds <-read.csv("data/small_diamonds.csv", header = T,
                          stringsAsFactors = T)
diamond_compare <-read.csv('data/diamond_compare.csv', 
                           header =T, stringsAsFactors = T)
source('source_files/general.R')
```

# Our data today
The `diamonds` data set comes pre-loaded with the `ggplot2` library, but it is quite large.  To simplify a little, we will work with 400 randomly sampled diamonds from that data set; we store this information in the `small_diamonds` data frame.

Let's explore
```{r}
small_diamonds %>% head()

small_diamonds %>% str()
```

For data sets and functions that come pre-loaded into `R`, you can use a `?` to learn more about them.
```{r}
?diamonds
```

### Removing columns
For this session, we will focus on just a small subset of the variables in `small_diamonds`. The `select()` function from the `dplyr` package allows us to *select* a subset of columns from a larger data frame.  

Let's write some code together that creates a new data frame `diamonds2` that only contains the variables: `carat`, `cut`, `depth`, `price`.
```{r}
# First build the dplyr pipe and make sure you get the right data frame

# Once you know your code is working, do an assignment to the shortcut name diamonds2.  You can do forward assignment -> or backward assignment <-.

```

## Percentiles
Rough guide, my exam grade was better than 72\% of my classmates grades, I would say *I scored in the 72nd percentile*.  

* Notice you do not need to know my actual grade to say this; it is simply a measure of where I stand among the other exam grades

Often we want to reverse this situation: how high do I need to score in order to beat 85\% of my classmates? Said differently, *What is the 85th percentile score?*


### Logical vectors and percentiles
A logical vector is simply a vector of TRUE or FALSE values, but `R` uses them in a number of ways. In general, `R` will allow you to perform *quantitative* summaries on logical vectors, and it does this by treating:

* TRUE is the same as 1
* FALSE is the same as 0.

This has some fun implications!  Yes, I really think this is fun... why do you ask?

Run the following code and think about what it does:
```{r}
# first create a simple vector of numbers:
numbers<- c(1, 2, 3, 4, 5, 6, 7)
numbers
# What does this code do?
numbers < 4
```

<Explain in your own words>

Now let's see how `R` will treat this "logical vector", `numbers < 4`, if we try to `sum()` it:
```{r}
sum(numbers<4)
```

**Exercise:** What percentage of the diamond prices in `diamonds2` are above the mean price?
```{r}
# Here's a guide (not the only way though):

# Compute the mean, save it under mean_price

# Create a logical vector for price values above  mean_price

# Use the logical vector to find the total number of diamonds priced above the mean

# take that total number and divide by the total rows in the data frame (nrow() helps)
```

The mean, 3655.418, is actually the _______ percentile! Is that what we usually expect?


## The harder direction
Suppose we are interested in the following question instead:

* What is the 80th percentile for the `carat` measurement?  
* Said differently, how many carats would a diamond need to weight in order to be heavier than 80\% of the other diamonds in the data set?

We rely on the `quantile()` command to find these quickly.  To use this properly, you first put your data vector, then pass the percentile you want in next ***as a decimal***. Let's find the 80th percentile of the `carat` variable:
```{r}
#Place code here
```

Yes, we can use this as part of a `summarize()` table, so long as we only compute one percentile at a time. Try adding the 80th percentile to the table below.
```{r}
diamonds2 %>% 
    summarise("mean (carat)" = mean(carat),
              std.dev = sd(carat))
```


We can compute a bunch of percentiles at once by passing a vector of probabilities into the `probs = ` argument. For example, try putting `c(.1, .25, .5, .75, .9)` into that argument:
```{r}
# Place code here.
```

**Quartiles** are a fancy way of referring to the 25th, 50th, and 75th percentiles.  Specifically,

* The first quartile $Q_1$ is the same as the 25th percentile $P_{25}$
* The second quartile $Q_2$ is the same as the 50th percentile $P_{50}$ (also the median)
* The third quartile $Q_3$ is the same as the 75th percentile $P_{75}$

The `summary()` command is a great way to get the quartiles, as well as the min, max, and mean of the data set.  Try applying `summary()` to the 
```{r}
#  take your guess!
```

#### A new measure of variation
We define the *inter-quartile range* (IQR) to be the difference between the third and first quartiles: Q3 - Q1.  The `IQR()` function computes this quickly in R.

## Why do we ***care?***
Let's compare two diamond collections. For a private collection, and for the Smithsonian collection, we randomly select 50 diamonds from each.

Run the code below to compare the mean and standard deviation of the carat measurement for each collection.
```{r}
diamond_compare %>%
    group_by(collection) %>%
    summarise(mean = mean(carat),
              std.dev = sd(carat)
              ) %>%
    # Remember you can only use flip_tibble() inside our lessons.
    flip_tibble()
```

Now, add median and IQR (our "percentile-based" measure of center and spread).  What do you notice?

## Boxplots
Let's try to visualize the difference we are observing above.  One common visual that relies on "percentile" measurements is a boxplot.  

To make a boxplot in ggplot, we assign one axis aesthetic to the grouping variable (usually `x`), and the other to the quantitative variable (usually `y`). Then, we simply use the geometry `geom_boxplot()`. Try it below!

```{r}
# Try to write you code from scratch!
ggplot(data = diamond_compare,
       aes(x = collection,
           y = carat))+
    geom_boxplot()
```

Yikes! Our boxplot got smushed, but at least we are starting to see what the issue might be.... 

**Exercise**. How many diamonds weigh more than 2.5 carats?
```{r}
# nice review, can you write the code?

```


**Exercise**. Copy and paste your boxplot code above, but this time add the layer `coord_cartesian(ylim=c(0,2.5))`.  This will refocus the boxplot to only diamonds with weight between 0 and 2.5 carats.  What do you notice?
```{r}
ggplot(data = diamond_compare,
       aes(x = collection,
           y = carat))+
    geom_boxplot()+
    coord_cartesian(ylim=c(0,3))
```


### Challenges!

**Exercise**: 
Can you create histograms for each in a single plot?
```{r}
# write your code here...
```

Yikes, weird looking histograms. 

**Exercise**.
Try making the following modifications (yes, I had to play around to get this just right):

* use `bins = 200` and `color="white"` inside the histogram geometry
* add the layer `coord_cartesian(xlim=c(0,3))`.  (Notice `xlim` and not `ylim` here...)

```{r}
# copy and paste your code above, add the coord_cartesian() layer
ggplot(data = diamond_compare,
       aes(x = carat))+
    geom_histogram(bins = 200,color="white")+
    facet_wrap(vars(collection), nrow=2)+
    coord_cartesian(xlim=c(0,3))
```

**Exercise.** How much does the "outlier" diamond weigh?
```{r}
# Place your code here.
```

### Final thoughts
So what's going on with the sample from the Smithsonian collection? I ***hope*** you figure it out.
