---
title: "Hypothesis Testing For Differences!"
subtitle: "Two at a time please!"
author: "P. Lombardo"
output:
  html_document:
    df_print: paged
---

## Loading packages, etc.
```{r}
library(tidyverse)
library(patchwork)
# library(DescTools)
library(broom)
fish <- read.csv('data/fish.csv', header = T, stringsAsFactors = T)
```

## Introduction
We have already explore how to compare population means or population percentages between two populations using confidence intervals.  This technique relied on a version of the Central Limit Theorem applying the following statistics:

1. The difference of sample means from two populations: $\bar x_1 - \bar x_2$.
2. The different of sample percentages from two populations: $\hat p_1 - \hat p_2$.

The good news is that hypothesis testing and confidence intervals work on the same underlying theory: a sampling distribution having a normal shape with formulas for the parameters.

Remember that the parameters we estimate are always *differences*, so our hypotheses will be in terms of what we expect about $\mu_1 - \mu_2$  (or $p_1 - p_2$).

This guide will be helpful to remember:

* Two-tailed test:
    * $H_0:\mu_1 - \mu_2 = 0$
    * $H_1:\mu_1 - \mu_2 \neq 0$
    * Intuitive meaning if reject null: $H_0:\mu_1 = \mu_2$
    * `alternative = "two.sided"`

* Right-tailed test:
    * $H_0:\mu_1 - \mu_2 \leq 0$
    * $H_1:\mu_1 - \mu_2 > 0$
    * Intuitive meaning if reject null: $\mu_1 > \mu_2$
    * `alternative = "greater"`
    
* left-tailed test:
    * $H_0:\mu_1 - \mu_2 \geq 0$
    * $H_1:\mu_1 - \mu_2 < 0$
    * Intuitive meaning if reject null: $\mu_1 < \mu_2$
    * `alternative = "less"`



## `ToothGrowth` revisited

Let's work once more with the `ToothGrowth` data set and test the claim that OJ leads to the same or less tooth growth over VC (vitamin C), when we consider only a dosage of 1.0 mg/day.

0. How should we prepare our data?
```{r}
#Place code here
```

1. Who is group 1? Let's run a levels command to see which comes first.

```{r}
#place code here.
```

*What difference are we making a claim about? (Make the order clear.)*


*What are the Null and Alternative Hypotheses?*

* $H_0:...$
* $H_1:...$

*What type of test is this? How should we set `alternative = ...`*


2. Assumptions of the hypothesis test:

Here are the assumptions for the hypothesis test. Notice we have an additional assumption of "equal variances."

* Either each group has at least 30 observations, or it is safe to assume we are sampling from two normal populations.
```{r}
# check sample sizes...
```

* The standard deviations are not wildly different (Three-fold-rule); let's use `group_by()` and `summarize()`:
```{r}
#place code here.
```

(*Note:* This standard deviation guideline is not strictly necessary with `t.test()` when `var.equal = FALSE`...)

3. Let's develop our shortcut code:

* Recall this is a right-tailed test, so `alternative = "greater"`
* We have two variables now, a response and an explanatory variable.
    * Response: ...
    * Explanatory: ...
    * `Response ~ Explanatory`


```{r}
#place code here...
```


4. Conclusion, assuming a significance level of 1\%:

....


## Returning to Fish
Now let's practice on our own.

0. Create a boxplot comparing all four species of tuna.

*Show Work Below:*


```{r}
#place code here.
```


Pick two species of tuna, and compare a difference in their population mean weights *using a hypothesis test.*

*Show Work Below:* (Follow the same steps as above.)


```{r}
#place code here.
```

